
---
title: "Association rules"
date: "2023"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())
```

# Load the required libraries and the database

```{r}
library(arules)
library("arulesViz")
library(dplyr)
library(ggplot2)
library(Hmisc)

dd <- read.table("2.Bank_India_preprocessed_data.csv",header=T, stringsAsFactors=TRUE, sep=",")

#Selecting categorical variables
dcat<-dd[,sapply(dd, is.factor)]
```


# 1) Transform into a transactional database 
The transaction database will have the same number of rows as the original but the same number of levels as categorical values we have.

As we already knew, our target and other categories of the database
are unbalanced.

```{r}
dtrans<-as(dcat, "transactions")

foo<-function(x){length(levels(x))}
sum(sapply(dcat, foo))

#Checking our transactional database
dtrans
inspect(head(dtrans,10))
summary(dtrans)
```


```{r}
itemFrequencyPlot(dtrans, topN=10, main ="Top 10 frequent itemsets", ylim = c(0,1))
itemFrequencyPlot(dtrans, support = 0.5, main ="Top 10 frequent itemsets", ylim = c(0,1))
```

Then, we did the same analysis with ECLAT

```{r}
ecl = eclat(dtrans)

size_eclat = size(ecl)

data.frame(size_eclat) %>%
  ggplot(aes(x = size_eclat)) +
  geom_histogram() +
  labs(title = "ECLAT: Distribution of the size of the transactions",
       x = "Size", y = "Frequency") + 
  theme_bw() + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 8))
```

# 2) General rules 

First, we wanted to discover all the important rules in the database. To do that, we define a low support (0.01) and a big confidence (0.7) and remove the empty rules by setting minlen=2. Then, we filter by lift to consider only good rules. We also tried also to establish an smaller value of support (0.01), but it does not provide any additional good rules

```{r}
rules <- apriori(dtrans, parameter =
                      list(support = 0.01, confidence = 0.7, minlen=2, maxlen = 11))
good.rules <- subset(rules, subset = lift > 1)

# Remove redundant rules
nonredundant <- good.rules[!is.redundant(good.rules),]
summary(nonredundant)

inspect(head(nonredundant,n=10, by="lift"))

length(good.rules)
length(nonredundant)
good.rules <- nonredundant
```

Here we demostrate that 0.01 is a good value for support for our case, as it provides
almost the same number of rules than a lower value

```{r}
df.rules <- DATAFRAME(good.rules)
plot(sort(df.rules$support, decreasing = TRUE), type = "l", main = 'Rules and support', ylab = 'Minimum support')
minor.tick(ny = 2, tick.ratio = 0.5)
abline(h=0.01, col="red", lty = 3, lwd = 2)
```


## Analysis of the rules

We consider that a rule is good if it has a lift => 3 and is not redundant. With this criteria there are 85 rules that we need to analyse.

We see that highest lift rules (lift = 17) are trivial, since they relate people with the same occupation and job_type. This was expected because these variables tell a
similar information.

The most relevant rule is that the set of pensioners is exactly the set
of people with job type unknown.

```{r}
# Note that setting an smaller support does not provide more good rules
rules <- apriori(dtrans, parameter =
                      list(support = 0.01, confidence = 0.7, minlen=2, maxlen = 11))

good.rules <- subset(rules, subset = lift > 3)

nonredundant <- good.rules[!is.redundant(good.rules),]

# Save the rules
# write.csv(DATAFRAME(nonredundant))
```

```{r}
# Explore and plot the rules
library(colourvalues)

length(nonredundant)
# inspect(head(nonredundant, n=80, by="lift"))

plot(nonredundant, measure = c("support", "lift"), shading = "confidence", colors = c( 'red', 'blue', 'green'), interactive = FALSE)
```

Here we analyse the case pensioners and Jobtype_Unknown and discovered that 
all pensioners reported "Jobtype_Unknown". 

```{r}
pensioners = which(dcat$job_stat=="Pensioner")
unk_jobtype = which(dcat$job_type=="Jobtype_Unknown")
setdiff(pensioners, unk_jobtype)
setdiff(unk_jobtype, pensioners)

inspect(sort(good.rules, by="lift")[9:30,])
# Many rules associate people with unknown occupation and pensioners. Let's
# look at their setdiff()

unk_occupation <- which(dcat$occupation == "Occupation_Unknown")
setdiff(pensioners, unk_occupation)
setdiff(unk_occupation, pensioners)
length(setdiff(unk_occupation, pensioners))
# All pensioners declared their occupation as unknown. However, 675 individuals
# didn't tell their occupation and weren't pensioners.
```

Now we will analyze the category unk_occupation to try to discover rules 
about the unknowns in Jobtype.

```{r}
unk.occup.not.pensioner <- subset(good.rules, subset =
                                    lhs %in% "occupation=Occupation_Unknown" & 
                                    !(rhs %in% "job_stat=Pensioner") &
                                    !(rhs %in% "job_type=Jobtype_Unknown"))
inspect(head(unk.occup.not.pensioner, n=10, by="lift"))
unk.occup.not.pensioner2 <- subset(good.rules, subset =
                                    lhs %in% "occupation=Occupation_Unknown" & 
                                    !(rhs %in% "job_stat=Pensioner") &
                                    !(rhs %in% "job_type=Jobtype_Unknown") &
                                    size(lhs) <= 1)
inspect(head(unk.occup.not.pensioner2, n=10, by="lift"))

# We'll try now to reverse the rules
unk.occup.not.pensioner3 <- subset(good.rules, subset =
                                    rhs %in% "occupation=Occupation_Unknown" & 
                                    !(lhs %in% "job_stat=Pensioner") &
                                    !(lhs %in% "job_type=Jobtype_Unknown") &
                                     size(lhs) <= 5)
inspect(head(unk.occup.not.pensioner2, n=10, by="lift"))

unk.occup.not.pensioner4 <- subset(good.rules, subset =
                                     rhs %in% "occupation=Occupation_Unknown" & 
                                     !(lhs %in% "job_stat=Pensioner") &
                                     !(lhs %in% "job_type=Jobtype_Unknown") &
                                     size(lhs) <= 4)
inspect(head(unk.occup.not.pensioner4, n=10, by="lift"))
# We can't describe in a confident and simple way the individuals with unknown
# occupation who aren't pensioners
```

Now, we will study the simplest rules. To find them, we reduce the limit of
the lift to 1.

```{r}
# We'll study now the simplest rules
rules2 <- apriori(dtrans, parameter =
                   list(support = 0.01, confidence = 0.7, minlen=2, maxlen = 2))
good.rules2 <- subset(rules2, subset = lift > 1)

inspect(head(good.rules2, n=10, by = "lift"))
# Highest lift rules are trivial, since they relate people with the same
# occupation and job_type. This was expected because these variables tell a
# similar information

# Before doing any feature selection, let us remove redundant rules
nonredundant2 <- good.rules2[!is.redundant(good.rules2),]
summary(nonredundant2)
inspect(head(nonredundant2,n=10, by="lift"))

length(good.rules2)
length(nonredundant2)
good.rules2 <- nonredundant2

inspect(head(good.rules2, n=25, by="lift"))
# With these simple rules, we are able to describe better our individuals by
# gender and the jobs that make job_stat=Working up. Other categories we understand
# better now are car=N and studies = Secondary education.
# However, these rules are not relevant for our goals in this analysis. Moreover,
# none of them have a big lift value.
```

Lastly, we will plot the best rules that we have discovered

```{r}
top10 <- sort(nonredundant, by="lift")[1:10]

plot(top10, method="grouped", shading = "lift")
```


# Rules related to the target

Finally, we decided to study our target starting with the category "payed" by putting
this category in the right and left hand side. 

```{r}
rules <- apriori(dtrans, parameter =
                      list(support = 0.01, confidence = 0.7, minlen=2, maxlen = 11))

good.rules <- subset(rules, subset = lift > 1)

payedRules <- sort(subset(good.rules, subset = rhs %in% "target=payed"), by = "lift")
summary(payedRules)

inspect(head(payedRules, by="lift"))
```


```{r}
# Plot the rules
plot(payedRules, measure=c("support", "confidence"), shading="lift",
     main = 'Association rules (8266) with payers in the rhs') + theme(legend.position = "bottom")
```


```{r}
payedRules2 <- sort(subset(good.rules, subset = (lhs %in% "target=payed" & lift > 3)), by = "lift")
inspect(head(payedRules2, n=10, by="lift"))

# We can try to filter, removing "jobtype unknown" and "medicine" that are trivial rules

payedRules2 <- sort(subset(good.rules, subset = (lhs %in% "target=payed" & !(rhs %in% "job_type=Medicine") & !(rhs %in% "job_type=Jobtype_Unknown") & !(lhs %in% "job_type=Jobtype_Unknown") & lift > 3), by = "lift"))
inspect(head(payedRules2, n=10, by="lift"))
```

```{r}
plot(payedRules2, measure=c("support", "confidence"), shading="lift",
     color = c('blue', 'lightblue'), main = 'Association rules (3570) with payers in the lhs') + theme(legend.position = "bottom")
```

Now, we will predict the people that did not pay
```{r}
overdueRules <- sort(subset(good.rules, subset = rhs %in% "target=overdue"), by = "lift")
summary(overdueRules)
```

As there are no rules on the rhs we decided to reduce the support to try to find some

```{r}
# Note that setting an smaller support does not provide more good rules
rules2 <- apriori(dtrans, parameter =
                      list(support = 0.001, confidence = 0.7, minlen=2, maxlen = 11))
good.rules2 <- subset(rules2, subset = lift > 3)
nonredundant2 <- good.rules2[!is.redundant(good.rules2),]

length(nonredundant2)

overdueRules2 <- subset(nonredundant2, subset =
                                     (rhs %in% "target=overdue"))

# Note: even if there are good rules the support is too low, so if they exist,
# are very rare and, then, not useful to predict the people that did not pay
```


Later, we checked for the characteristics on the left part

```{r}
overdueRules3 <- sort(subset(good.rules, subset = (lhs %in% "target=overdue" & 
                             lift > 2), by = "lift"))
inspect(head(overdueRules3, by="lift"))

#plot(overdueRules3, measure = c("support", "lift"), shading = "confidence")
plot(overdueRules3, method="graph", shading = "lift") + theme(plot.background = element_rect(fill='transparent', color=NA))
```





