---
title: "Preprocessing MVA-11-BANK-INDIA"
author: "Adrià Casanova Lloveras"
date: "2023-10-26"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())
```


# Library importation
```{r Libraries, include=FALSE}
library(tidyverse)
library(memisc)
library(dplyr) 
library(recipes)
library(stats)
library(caret)
library(class)
library(MASS)
library(mvnmle)
library(devtools)
library(ggplot2)
library(misty)
library(mice)
source("LittleMCAR function.R")  # Import LittleMCAR function from the folder.
```


##### Data Overview #####

# Read data
```{r}
data <- read.csv("1. Bank-India-raw-data.csv")

summary(data)
```


##### Preprocessing - Basic Transformation #####

# Rename the features to a naunced name
```{r}
names(data) <- c("id", "target", "contract", "gender", "car", "n_child",
                 "income", "credit", "loan", "price", "job_stat", "studies",
                 "family", "house", "age", "job_duration", "occupation",
                 "job_type", "n_enquiries", "companion")
names(data)
```

# Rename the categorical feature in a more concise name, and group them into bigger sets.
```{r}
data$job_type <- sub(":.*", "", data$job_type)
data$companion[data$companion %in% c("Other_A", "Other_B", "Group_people")] <- "Other_companion"
data$family[data$family %in% c("Civil marriage")] <- "Married"
data$occupation[data$occupation %in% c("Accountants", "Secretaries","HR staff") ] <- "Administrative Staff"
data$occupation[data$occupation %in% c("High skill tech staff", "IT staff") ] <- "Tech Staff"
data$occupation[data$occupation %in% c("Low-skill Laborers", "Laborers") ] <- "Laborers"
data$occupation[data$occupation %in% c("Cleaning staff", "Chef","Drivers","Waiters") ] <- "Service Staff"

data$studies[data$studies %in% c("Incomplete higher","Lower secondary")] <- "Low education"
data$studies[data$studies %in% c("Secondary / secondary special")] <- "Secondary education"
data$studies[data$studies %in% c("Higher edu." = "Higher education","Academic degree")] <- "Higher education"

data$job_type[data$job_type %in% c("Business Entity Type 1", "Business Entity Type 2", "Business Entity Type 3")] <- "Business"
data$job_type[data$job_type %in% c("Advertising", "Bank", "Business") ] <- "Business"
data$job_type[data$job_type %in% c("Culture", "Electricity", "Emergency", "Hotel", "Industry","Legal Services", "Services", "Restaurant", "School", "Kindergarten", "Other") ] <- "Culture and Services"
data$job_type[data$job_type %in% c("Government", "Military", "Security Ministries", "Police", "Postal") ] <- "Government and Military"
data$job_type[data$job_type %in% c("Realtor", "Trade", "Telecom") ] <- "Real Estate and Trade"
data$companion[data$companion %in% c("Children", "Family")] <- "Family"


# recode shorter name some modalities
data$job_stat <- fct_recode(data$job_stat, "Commer. Assoc." = "Commercial associate")
data$family <- fct_recode(data$family,  "single" = "Single / not married", "divorce" = "Separated")
data$house <- fct_recode(data$house, "apartment" = "House / apartment", "Municipal apart." = "Municipal apartment", "Rented apart."="Rented apartment", "Office apart."="Office apartment", "Co-op apart."="Co-op apartment")
data$occupation <- fct_recode(data$occupation, "Medic stf" = "Medicine staff", "Chef" = "Cooking staff", "Security" = "Security staff", "Waiters" = "Waiters/barmen staff", "Private ser." = "Private service staff")
data$companion <- fct_recode(data$companion, "Partner" = "Spouse, partner", "Unaccompan." = "Unaccompanied", "Group_people" = "Group of people")

```

# Transform the numerical & characters into factor.
```{r}
# convert numerical target (0,1) -> as factor ("payed", "overdue")
data$target <- factor(data$target, levels = c(0,1), labels = c("payed","overdue"))

# convert character feature (such as gender, car or contract type) -> as factor:
char_cols <- which(sapply(data, is.character))
data[, char_cols] <- lapply(data[, char_cols], as.factor)


# Review if the transformation has executed properly
str(data)
summary(data)
```

# Transform variables 
# First, we have an age feature (age when they applied to the loan) that is a negative number and computed as the days spent in their lifetime until the moment they apply for the loan. We want to transform it into years, for a more comprehensive purpose.
```{r}
data$age<-floor(abs(data$age) / 365)
```

# Second, we transform the job_duration variable (#days they worked when they applied to the loan in the current work). Also calculated with days. We think it is more understandable to see the years. 

# When we have a glimpse of the values, we can see an error in this feature. There is a value of 365243 days working, which cleary is a wrong value as it's equivalent to 1000 years worked. This value is recorded when the job_type equal to XNA, which means the applicant does not have any work at the time of application.
# So we also rewrite the missing values of the factor variables as "Unknown". 
```{r}
# data$job_duration
data$job_duration[which(data$job_duration == 365243)] <- NA
data$job_duration <- (abs(data$job_duration) / 365) # job duration in years
data$job_type <- as.character(data$job_type)
data$job_type[which(data$job_type == "XNA")] <- "Unknown"
data$job_type <- as.factor(data$job_type)

data$gender <- as.character(data$gender)
data$gender[which(data$gender == "XNA")] <- "Unknown"
data$gender <- as.factor(data$gender)

data$occupation <- as.character(data$occupation)
data$occupation[which(data$occupation == "")] <- "Unknown"
data$occupation <- as.factor(data$occupation)

data$companion <- as.character(data$companion)
data$companion[which(data$companion == "")] <- "Unknown"
data$companion <- as.factor(data$companion)

```

# Check if the values and NA's are correctly saved
# We assign to the NA's in factor variables as "Unknown". 
```{r}
summary(data)
```

# The variables that have (other) : occupation, job_type,companion: means that there are more classes. Let's explore them: 
```{r}
names(table(data$occupation))
names(table(data$job_type))
names(table(data$companion))
```

# We calculated the percentage of missing values which is NA's + Unknowns ratio
```{r}
missing.values.data <-  as.data.frame(skimr::skim(data))
na_perc = sum(missing.values.data$n_missing) / (ncol(data) * nrow(data)) * 100
unknown_count = sum(data$job_type == "Unknown", na.rm = T) + sum(data$gender == "Unknown") +
  sum(data$occupation == "Unknown") + sum(data$companion == "Unknown")
unknown_perc = unknown_count / (ncol(data) * nrow(data)) * 100
missing_perc = na_perc + unknown_perc
```

# Attention, there is only one numerical variable with missing values: 'n_enquiries'. Percentage of missing values:
```{r}
689/nrow(data)*100
```

# Run the little test to see if the missing value are MCAR, MAR or MNAR type
```{r}
little.test <- LittleMCAR(data)
```
```{r}
little.test$chi.square
```

# Data is NOT missing completely at random as p-value is 0.

# We assume that is at RANDOM : then we have to impute
```{r}
little.test$p.value
```
```{r}
little.test$df
```
```{r}
little.test$missing.patterns
```
```{r}
little.test$amount.missing
```


######  Preprocessing - Data Imputation #######

# Only select numerical variables
```{r}
numerical_data <- data %>%
  select_if(is.numeric)

md.pattern(numerical_data)
```

# Use MICE to impute the data
```{r}
# Look the NA's with VIM packages
mice_plot <- aggr(numerical_data, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(numerical_data), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```
```{r}
# use mice to impute the missing numerical values
imputed_Data <- mice(numerical_data, m=5, maxit = 50, method = 'pmm', seed = 500)
summary(imputed_Data)
```
```{r}
# inspect quality of imputations
stripplot(imputed_Data, price, pch = 19, xlab = "Imputation number")
imputed_Data$imp$price

# get complete data ( 3nd out of 5)
completeData <- mice::complete(imputed_Data, 3)
```
```{r}
stripplot(imputed_Data, job_duration, pch = 19, xlab = "Imputation number")
```
```{r}
stripplot(imputed_Data, n_enquiries, pch = 19)
```
```{r}
stripplot(imputed_Data, pch = 10, cex = 1.2)
```
```{r}
densityplot(imputed_Data)
```

# KNN for numerical imputations
```{r}
knn_imputed <- kNN(numerical_data)
```
```{r}
marginplot(knn_imputed[, c("price", "job_duration")])
```

# MIMI function for numerical imputations
```{r, include=FALSE}
#assume missings represented with NA
uncompleteVar<-function(vector){any(is.na(vector))}

Mode <- function(x) 
{
  x<-as.factor(x)
  maxV<-which.max(table(x))
  return(levels(x)[maxV])
}

MiMMi <- function(data, priork=-1)
{
  #Identify columns without missings
  colsMiss<-which(sapply(data, uncompleteVar))
  if(length(colsMiss)==0){
    print("Non missing values found")
    out<-dd
    }else{
    K<-dim(data)[2]
    colsNoMiss<-setdiff(c(1:K),as.vector(colsMiss))
  
    #cluster with complete data
    dissimMatrix <- daisy(data[,colsNoMiss], metric = "gower", stand=TRUE)
    distMatrix<-dissimMatrix^2
  
    hcdata<-hclust(distMatrix, method = "ward.D2")
    plot(hcdata)
    nk<-2
    if(priork==-1){
      print("WARNING: See the dendrogramm and ZOOM if required")
      print("and enter a high number of clusters")
      nk<-readline("(must be a positive integer). k: ")
      nk<-as.integer(nk)
    }else{nk<-priork}
  
    partition<-cutree(hcdata, nk)

    CompleteData<-data
    #nomes cal per tenir tra?a de com s'ha fet la substituci?
    newCol<-K+1
    CompleteData[,newCol]<-partition
    names(CompleteData)[newCol]<-"ClassAux"
  
    setOfClasses<-as.numeric(levels(as.factor(partition)))
    imputationTable<-data.frame(row.names=setOfClasses)
    p<-1
  
    for(k in colsMiss)
    {
       #Files amb valors utils
       rowsWithFullValues<-!is.na(CompleteData[,k])
    
       #calcular valors d'imputacio
       if(is.numeric(CompleteData[,k]))
       {
          imputingValues<-aggregate(CompleteData[rowsWithFullValues,k], by=list(partition[rowsWithFullValues]), FUN=mean)
       }else{
          imputingValues<-aggregate(CompleteData[rowsWithFullValues,k], by=list(partition[rowsWithFullValues]), FUN=Mode)
       }
    
       #Impute
    
       for(c in setOfClasses)
       {
          CompleteData[is.na(CompleteData[,k]) & partition==c,k]<-imputingValues[c,2]
       }
    
       #Imputation Table
       imputationTable[,p]<-imputingValues[,2]
       names(imputationTable)[p]<-names(data)[k]
       p<-p+1
    }
    
    rownames(imputationTable)<-paste0("c", 1:nk)
    out<-new.env()
    out$imputedData<-CompleteData
    out$imputation<-imputationTable
  }
  return(out)
}
```
```{r}
mimi_imputed<-MiMMi(numerical_data)
```
```{r}
mimi_imputed$imputation
```
```{r}
mimi_imputed$imputedData
```

# Density Plot
```{r}
densityplot(data$price, col = "blue", main = "Density Plot of Price")
densityplot(mimi_imputed$imputedData$price, col = "blue", add = TRUE)
```
```{r}
densityplot(mimi_imputed$imputedData$job_duration, col = "blue", add = TRUE)
densityplot(data$job_duration, col = "blue", add = TRUE)

```
```{r}
densityplot(mimi_imputed$imputedData$n_enquiries, col = "blue", add = TRUE)
densityplot(data$n_enquiries, col = "blue", add = TRUE)
```
```{r}
densityplot(completeData$job_duration, col = "blue", add = TRUE)
densityplot(data$job_duration, col = "blue", add = TRUE)
```
```{r}
densityplot(completeData$n_enquiries, col = "blue", add = TRUE)
densityplot(data$n_enquiries, col = "blue", add = TRUE)
```
```{r}
densityplot(completeData$price, col = "blue", add = TRUE)
densityplot(data$price, col = "blue", add = TRUE)
```

# Based on the performances of each imputation method that we had for our dataset, We decided to use MICE for numerical imputations
```{r}
columns_to_replace <- c("price", "n_enquiries", "job_duration")
data[columns_to_replace] <- completeData[columns_to_replace]
head(data)
```
