---
title: "EDA"
output: html_document
date: "2023-10-28"
---

First we import the database before and after the imputation to perform
a EDA analysis with SmartDEA library

```{r}
library(SmartEDA) 

db_original = read.csv("Bank-India-raw-data.csv")
db_imputed = read.csv("Bank-India-preprocessed-data.csv")

ExpReport(db_original, op_file = 'SmartEDA_original.html4')
ExpReport(db_imputed, op_file = 'SmartEDA_imputed.html4')

```

We can also do the analysis for the two groups of the target variable

```{r}
# Find all rows where people did not pay the rend

debtors = db_imputed[which(db_imputed$target == "overdue"),]
payers = db_imputed[which(db_imputed$target == "payed"),]

barplot(debtors$companion)
plot(payers$job_type)

ExpReport(debtors, op_file = 'SmartEDA_debtors.html')
ExpReport(payers, op_file = 'SmartEDA_payers.html')
```

As distributions seems not-normal we can do several normality tests to check it.
The null hypotesis is "normality"

```{r}
shapiro.test(db_imputed$age)

ks.test(db_imputed$loan, "pnorm")

#this can be done for all numerical variables, the result is the same
```

Lets try to compare the distribution of a variable with a normal distribution
to check the tests visually


```{r}
a = plot(density(log(db_imputed$price)), col = "blue", lwd = 2,
     main = 'Comparison of a normal distribution (red) and "price" (blue)',
     xlab = 'Price') # Densidad
x = log(db_imputed$price)
x2 <- seq(min(x), max(x), length = 40)
f <- dnorm(x2, mean(x), sd(x))
lines(x2, f, col = "red", lwd = 3, type = 'l') # Normal
```
# Correlation of variables
For numerical variables, we can use the "pearson" coefficient to detect if
two variables are correlated

```{r}
library(corrplot)

numeriques <- which(sapply(db_imputed,is.numeric))

corrplot(cor(db_imputed[,numeriques]),
         method = "number",
         type = "upper" # show only upper side
)

```

Using the DEA reports, we see that it may be interesting to plot income and loan

```{r}
plot(db_imputed$income, db_imputed$loan, col=c("black", "red") [factor(db_imputed$target)],
     xlab = 'Income', ylab = 'Loan')
legend("topright", legend=c("Payed", "Overdue"), col=c("black", "red"), pch = 1)
```







