```{r}
library(arules)
library("arulesViz")
```
```{r}

data("Epub")
data("Groceries")
data("Adult")

#setwd your own working directory to import the database
#How to work with associations rules with data matrix

dd <- read.table("credscoClean.csv",header=T, stringsAsFactors=TRUE, sep=";");

#Selecting categorical variables
dcat<-dd[,sapply(dd, is.factor)]


#1) Transforming our table as a transactional database 
# Rows: the same number of rows
# Items: the same number of levels as categorical variables

dtrans<- as(dcat, "transactions")

#Checking levels
length(levels(dcat$Dictamen))
#[1] 2
length(levels(dcat$Vivienda))
#[1] 7
length(levels(dcat$Estado.civil))
#[1] 6
length(levels(dcat$Registros))
#[1] 2
length(levels(dcat$Tipo.trabajo))
#[1] 5

foo<-function(x){length(levels(x))}
sum(sapply(dcat, foo))

#Checking our transactional database
# Because we have 5 categorical it makes sense to have cases of size 1
dtrans
inspect(head(dtrans,10))
summary(dtrans)
dim(dtrans)
plot(size(dtrans))
plot(size(Groceries))
transactionInfo(Epub[1:10])

inspect(dtrans[1:10])

# Test and error. Tell me from the transactional database, with a support all the common 
# items

# If we see that there are little rules and we want more, then we need to reduce the 
# support
itemFrequencyPlot(Groceries, support=0.1, cex.names = 1)
itemFrequencyPlot(Groceries, support=0.05, cex.names = 1)
itemFrequencyPlot(Groceries, topN=5, cex.names = 1)
itemFrequencyPlot(dtrans, topN=5, cex.names = 1)
itemFrequencyPlot(dtrans, support=0.1, cex.names = 1)

# We decided to stop. My minimal stop if [].

#Apriori: Start analyzing 
?apriori

#minlen : minimum number of items

# There are 46 rules if we set confidence to 50%
# The confidence is bad, only 50% is not enough. So better increase it to 80%
rulesDtrans <- apriori(dtrans, parameter = list(support = 0.1, confidence = 0.5,  minlen=2))
rulesEpub <- apriori(Epub, parameter = list(sup = 0.1, conf = 0.5, target="rules", minlen=1))
rulesGroceries <- apriori(Groceries, parameter = list(sup = 0.01, conf = 0.5,  minlen=2))
rulesGroceries
summary(rulesGroceries)
inspect(rulesGroceries)
summary(rulesDtrans)
inspect(rulesDtrans)
items<-apriori(dtrans, parameter = list(sup = 0.2,  target="frequent itemsets", minlen=2))
inspect(items)

# Increase it to 80%
# Min len is to get associations of at least 1 element but is useless
# because this means that the output implies empty
# [LSR] = 0 ; A implies nothing if len=1

# Min len is at least 2
rulesDtrans <- apriori(dtrans, parameter = list(support = 0.4, confidence = 0.8,  minlen=2))

# Important command 
# use "lift". If the 3rt quantile is not even 2 all the rules are bad. We need to
# repeat the analysis
inspect(rulesDtrans)

# Rank the rules by the best lift
inspect(head(rulesDtrans,n=3, by="lift"))

data("Groceries")
mbarules<-apriori(Groceries, parameter = list (support=0.01, confidence=0.1, maxlen = 2))
inspect(head(mbarules,n=10, by="lift"))
inspect(head(mbarules,n=10, by="confidence"))

# See only the most frequent elements
mbaitemsets<-apriori(Groceries, parameter = list (support=0.01, confidence=0.1, maxlen = 2, target="frequent"))
inspect(head(mbaitemsets,n=10, by="support"))

#Searching for redundant rules

#
subset.matrix <- is.subset(rulesDtrans,rulesDtrans,sparse = F)
subset.matrix[lower.tri(subset.matrix,diag=T)] <- NA
redundant <- colSums(subset.matrix,na.rm=T)>=1
which(redundant)

#Removing redundant rules

rules.pruned <- rulesDtrans[!redundant]
rules.pruned <- sort(rules.pruned,by="lift")

# There is not a good rule. We need to reduce the numbers of confidence, do a better
# preprocessing, etc
inspect(rules.pruned)

# Redundant rules
# Ex of redundant rules
# <milk> = <yogurt>
# <milk, cheese> = <yogurt>

# Dante
# is.redundant(rules)


#´Karina script, does the same but longer

# The second part of the script is to change the order, a change called ECLAT.
# ECLAT is better for large databases, more efficient. But results are
# the same if we use the normal database or the ECLAT version


###You can use is.redundant and similar instructions to improve your analysis
### See the links of examples (English and Spanish Version)
### These links offers a way to summarize your Analysis
#### ********** https://rpubs.com/Buczman/AssociationRules
####*           https://rpubs.com/Joaquin_AR/397172
####These links is very useful to filter rules
##################################### according to your requirements, check "FILTRADO
##################################### DE REGLAS





###
### Saving results
write(rules.pruned, file = "rules.csv", sep = ",", col.names = NA)


#ECLAT

eclatDTrans<-eclat(dtrans)
eclatDTrans<-eclat(dtrans, parameter = list(support=0.4, minlen=1, maxlen=10))
inspect(eclatDTrans)

## Select a subset of rules using partial matching on the items 
## in the right-hand-side and a quality measure


## Mine frequent itemsets with Eclat.
eclatDTrans <- eclat(dtrans, parameter = list(supp = 0.5))

## Display the 5 itemsets with the highest support.
orderedItemsets <- sort(eclatDTrans)
inspect(orderedItemsets)

top5 <- sort(eclatDTrans)[1:5]
inspect(top5)

## Get the itemsets as a list
as(items(top5), "list")

## Get the itemsets as a binary matrix
as(items(top5), "matrix")

## Get the itemsets as a sparse matrix, a ngCMatrix from package Matrix.
## Warning: for efficiency reasons, the ngCMatrix you get is transposed 
as(items(top5), "ngCMatrix")

###Visualizing Results
inspect(rulesDtrans)
plot(rulesDtrans, measure = c("support", "lift"), shading = "confidence")
#order == number of items inside the rules
plot(rulesDtrans, method = "two-key plot")
plot(rulesDtrans, method = "grouped")
plot(rulesDtrans, method = "paracoord")

###Check the suggested links to get a friendly and visual analysis of your results
```

# Usefull links
rpubs.com/Buczman/AssociationRules

0.1 is a large support, so we need to change it to have more results.

After this analysis they generate the rules by ussing ECLAT, because the 
database is big. We could have done it without it. 

There are some patterns that are trivial, if you buy butter and yogurt it
make sense that you will buy whole and milk. 

Indusction methot
You filter rules, only using the products that ends in "whole milk"

we can present the final results in a network

Tree maps are better than pie graphs.


rpubs.com/Joaquin_AR/397172

Usefull because it tells us how to filter the rules


Use the basic script to detect rules and the links to plot them




