---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r Libraries, include=FALSE}
library(dplyr)
library(tidyverse)
library(skimr)
library(VIM)
library(mice)
library(cluster)
require(StatMatch)
library(ggplot2)
```

# Context
# Read data
```{r}
data <- read.csv("Bank-India-data.csv")
```
# Data Exploratory
# Preprocessing
### Rename the features to a understandable name
```{r}
names(data)
names(data) <- c("id", "target", "contract", "gender", "car", "n_child",
                 "income", "credit", "loan", "price", "job_stat", "studies",
                 "family", "house", "age", "job_duration", "occupation",
                 "job_type", "n_enquiries", "companion")
names(data)
```

### Assign proper type
```{r}
# convert  numerical-> as factor
data$target <- factor(data$target, levels = c(0,1), labels = c("payed","overdue"))
# convert character -> as factor:
char_cols <- which(sapply(data, is.character))
data[, char_cols] <- lapply(data[, char_cols], as.factor) 
str(data)
summary(data)
```


### Rename the Categorical feature in a shorter name.
```{r}
data$job_stat <- fct_recode(data$job_stat, "Commer. Assoc." = "Commercial associate")
data$family <- fct_recode(data$family,  "single" = "Single / not married", "divorce" = "Separated")
data$house <- fct_recode(data$house, "apartment" = "House / apartment", "Municipal apart." = "Municipal apartment", "Rented apart."="Rented apartment", "Office apart."="Office apartment", "Co-op apart."="Co-op apartment")
data$occupation <- fct_recode(data$occupation, "High-tech stf" = "High skill tech staff", "Medic stf" = "Medicine staff", "Chef" = "Cooking staff", "Security" = "Security staff", "Waiters" = "Waiters/barmen staff", "Low-skill labor." = "Low-skill Laborers", "Private ser." = "Private service staff")
data$companion <- fct_recode(data$companion, "Partner" = "Spouse, partner", "Unaccompan." = "Unaccompanied", "Group_people" = "Group of people")
data$studies <- fct_recode(data$studies, "Secondary" = "Secondary / secondary special", "Higher edu." = "Higher education")

```


### Check if it's possible to colapse the organization type into a single categoric variable.
### In the organization type, there is 4 variable with same name, but differente type. e.g. for Industry, there is type 1 to 12. Maybe we can resume it.
### For this purpose, I'll try to do the analysis the contingency table

```{r}
industry <- paste("Industry: type", 1:12)

job_industry <- data %>% filter(data$job_type %in% industry)
job_industry <- droplevels(job_industry)


con1 <- table(job_industry$job_type, job_industry$target)  
con1

mosaicplot(con1)  

chisq.test(con1)

```




### Transform variables 
First, we have an age variable (age when they applied to the loan) that is a negative number and computed as the days spent in their lifetime, we want to save the age they had when they applied to the loan in years. We created the new variable transformed_age. 
```{r}
data$age<-floor(abs(data$age) / 365)
```

Second, we transform the job_duration variable (#days they worked when they applied to the loan). Also calculated with days. We think it is more understandable to see the years. 

When we have a glimpse of the values, we can see an error in the data set. The value 365243 clearly is a wrong value because it is equivalent to 1000 years worked and it is found in many records. We will transform it to a missing value. 
We also rewrite the missing values of the factor variables as "Unknown". 
```{r}
# data$job_duration
data$job_duration[which(data$job_duration == 365243)] <- NA
data$job_duration <- (abs(data$job_duration) / 365) # job duration in years
data$job_type <- as.character(data$job_type)
data$job_type[which(data$job_type == "XNA")] <- "Unknown"
data$job_type <- as.factor(data$job_type)

data$gender <- as.character(data$gender)
data$gender[which(data$gender == "XNA")] <- "Unknown"
data$gender <- as.factor(data$gender)

data$occupation <- as.character(data$occupation)
data$occupation[which(data$occupation == "")] <- "Unknown"
data$occupation <- as.factor(data$occupation)

data$companion <- as.character(data$companion)
data$companion[which(data$companion == "")] <- "Unknown"
data$companion <- as.factor(data$companion)

```

### Check if the values and NA's are correctly saved
We assign to the NA's in factor variables as "Unknown". 
```{r}
summary(data)
```
The variables that have (other) : occupation, job_type,companion: means that there are more classes. Let's explore them: 

```{r}
names(table(data$occupation))
# What does the following code do?
# data <- data %>%
#  mutate_all(~replace(., . == "XNA", "Unknown"))
# print(unique(data$occupation))
names(table(data$job_type))
names(table(data$companion))
```
# NA's + Unknowns ratio
We calculated the percentage of missing values
```{r}
missing.values.data <-  as.data.frame(skimr::skim(data))
na_perc = sum(missing.values.data$n_missing) / (ncol(data) * nrow(data)) * 100
unknown_count = sum(data$job_type == "Unknown", na.rm = T) + sum(data$gender == "Unknown") +
  sum(data$occupation == "Unknown") + sum(data$companion == "Unknown")
unknown_perc = unknown_count / (ncol(data) * nrow(data)) * 100
missing_perc = na_perc + unknown_perc
```

Attention, there is only one numerical variable with missing values: 'n_enquiries'.
Percentage of missing values:
```{r}
689/nrow(data)*100
```


```{r}
library(ggplot2)
```

Find relation between age and n_child
```{r}
# df <- data.frame(data$n_child, age)

# Calculate the average age for each n_child value
avg_age <- tapply(data$age, data$n_child, mean)
regression_model <- lm(data$age ~ data$n_child, data = data)
```

```{r}
# Create the scatter plot
plot(names(avg_age), avg_age, xlab = "n_child", ylab = "Average Age", main = "Average Age vs. n_child")
abline(regression_model, col = "blue")
```

Plot avg(age) ~ n_child
```{r}
par(mfrow=c(1,1))
plot (data$n_child, data$age)
```

Regression model results
```{r}
regression_model
summary(regression_model)
```

Regression model plots
```{r}
par(mfrow=c(2,2))
plot(regression_model)
```
